Statement of Intent

### 2010-11-18 applications/nsf_fellowship/proposed_research.tex
Done with the application.
I made revisions on all of my essays (some minor, some major), and
I added a bib file, because I made some egregious changes to the
real bib files (to save space) that I didn't want to commit.
But I wanted to be able to remake these applications just as they are.

\textbf{Keywords.} \textit{Topic models, natural language processing,
information extraction}

@@

\textbf{Methods.}  Establishing that topic models can answer topic-based
queries the same way a human would requires two parts.  First one must show
that a topic found by a topic model can be accurately compared to a theme or
topic elicited from a human.  Second, one must show that an automated system
can correctly answer queries about topics in a corpus by returning the topics
from a topic model that match human answers to the same queries.

@@

\textbf{Validation.}  Once I have annotated data, I will be able to test the
validity of my hypothesis.  Determining whether a particular topic mathces a
human-defined topic should be straightforward with the annotations I will
gather.  I can test the words found by the topic model against the description
given by the humans, and I can see if the example documents given by the
annotators agree with the documents that have high proportions of the topic.

With methods for matching topics from a topic model with human-given topics, I
can complete the second part of my objective.  With topics obtained from humans
for a particular class, I can learn a topic model over the entire collection of
student ratings and see if automatic methods to rank the topics return the
correct topics.  One such method is to rank the topics by their entropy over
the documents for that class, which should rank highest the topics that were
used consistently about that class, matching the human-given topics and
answering the question, ``What were the main themes in comments of Chemistry
101?''  Other possible queries and methods to answer them are items that will
be explored in the course of this research.

\textbf{Broader Impacts.}  This work will quantitatively establish the
usefulness of topic models in extracting thematic elements from a collection of
documents.  It will also validate the automatic ranking of topics in response
to a few classes of specific user queries.  Such work will give confidence to
anyone using a topic model to analyze a corpus, including intelligence analysts
looking at classified documents, university administrators looking at student
ratings, and companies reading online reviews about their products.  This work
also establishes a method and a set of data for topic models to be compared to
each other in order to determine which models best capture human thematic
judgments.

\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\noindent \small\textbf{References}\vspace{-2mm}\footnotesize}
\bibliography{../../../bib/lda/bib}

>>>>

\textbf{Keywords.} \textit{Topic models, natural language processing,
information extraction, search}

@@

\textbf{Methods.}  Establishing that topic models can answer topic-based
queries the same way a human would involves two parts.  First one must show
that a topic found by a topic model can be accurately compared to a theme or
topic elicited from a human.  Second, one must show that an automated system
can correctly answer queries about topics in a corpus by returning the topics
from a topic model that match human answers to the same queries.

@@

\textbf{Validation.}  With annotated data, I can test the validity of my
hypothesis.  Determining whether a particular topic matches a human-defined
topic is straightforward with the annotations I will gather.  I can test the
words found by the topic model against the description given by the humans, and
I can see if the example documents given by the annotators agree with the
documents that have high proportions of the topic according to the model.

Once I have methods for matching topics from a topic model with human-given
topics, I can complete the second part of my objective.  I will learn a topic
model over each collection of documents and see if automatic methods to rank
the topics return topics that match those that were given by humans.  One such
method is to rank the topics by their entropy over the documents for the
annotated class or restaurant.  This ranking should match the human-judged
themes and answer the question, ``What were the main themes in comments about
this class?'' Similar methods can give a broad overview of the topics in the
entire corpus, and other possible queries and automatic ranking methods to
answer them are items that will be explored in the course of this research,
including finding topics that are common or unique to two separate classes.

\textbf{Broader Impacts.}  This work will quantitatively establish the
usefulness of topic models in extracting thematic elements from a collection of
documents.  It will also validate the automatic ranking of topics in response
to a few types of specific user queries.  Such work will give confidence to
anyone using a topic model to analyze a corpus, including intelligence analysts
looking at classified documents, university administrators looking at student
ratings, and companies reading online reviews about their products.  This work
also establishes a method and a set of data for topic models to be compared to
each other in order to determine which models best capture human thematic
judgments.

\footnotesize
\bibliographystyle{plain}
\renewcommand\bibsection{\noindent \small\textbf{References}\vspace{-2mm}\footnotesize}
\bibliography{bib}
